---
title: "Sex Proportions Across Penguin Species"
author: "Yingzhu Pan"
date: "2026-09-24"
categories: [R, Data Analysis]
engine: knitr
execute:
  echo: true
---

## Introduction

Are the proportions of male and female penguins similar across different species?
I use R to compare the sex proportion across Adelie, Chinstrap, and Gentoo penguins.

Data source: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/),
collected and made available by Dr. Kristen Gorman and
Palmer Station Antarctica LTER. The data are available under
the [CC0 license](https://creativecommons.org/publicdomain/zero/1.0/).
I access the data through the R `palmerpenguins` package.

Original study: Gorman, K. B., Williams, T. D., and Fraser, W. R.
(2014). Ecological sexual dimorphism and environmental variability
within a community of Antarctic penguins (genus Pygoscelis).
[PLOS ONE, 9(3), e90081](https://doi.org/10.1371/journal.pone.0090081).

## Load and inspect the data

I load the packages and inspect the dataset. Package startup
messages are hidden to keep the output focused on the data.

```{r}
#| label: load-data-r
#| message: false

library(palmerpenguins)
library(dplyr)
library(tidyr)
library(ggplot2)

penguin_data <- penguins
head(penguin_data)
```

## Clean the data

I keep records with non-missing species and sex. Missing values in other columns 
will not affect this analysis.

```{r}
#| label: clean-data-r

penguin_data <- penguin_data %>%
  drop_na(species, sex) %>%
  select(species, sex)

cat("Original rows:", nrow(penguins), "\n")
cat("Rows kept:", nrow(penguin_data), "\n")
cat("Rows removed:", nrow(penguins) - nrow(penguin_data), "\n")

head(penguin_data)
```

## Calculate sex proportions
I count female and male penguins within each species, then calculate
percentages using the number of penguins with recorded sex in that
species as the denominator.

```{r}
#| label: sex-proportions

sex_summary <- penguin_data %>%
  group_by(species) %>%
  count(sex) %>%
  mutate(percent = round(n / sum(n) * 100,1)) %>%
  ungroup()

sex_summary
```


## Compare sex proportions visually

Each species has two bars showing the percentages of female
and male penguins with recorded sex.

```{r}
#| label: fig-sex-proportions
#| fig-cap: "Female and male percentages within each penguin species, excluding records with missing species or sex."
#| fig-width: 7
#| fig-height: 5

ggplot(sex_summary, aes(x = species, y = percent, fill = sex)) +
  geom_col(position = "dodge") +
  labs(x = "Species", y = "Percentage (%)", fill = "Sex")
```

## Conclusion

The proportions of female and male penguins are similar across
the three species in this sample. Adelie and Chinstrap penguins
have equal proportions of females and males, while Gentoo
penguins have a slightly higher proportion of males.

These results include only penguins with recorded species and
sex. They describe the observed sample and do not necessarily
represent the sex proportions of the entire penguin populations.

